---
title: "DSE6211 Preliminary Results"
author: "Nelson Tran"
date: "2024-02-12"
output:
  word_document: default
  pdf_document: default
---

Loading libraries that may be used.
```{r}
library(lubridate)
library(reticulate)
library(keras)
library(tensorflow)
library(tidymodels)
library(caret)
library(MESS)
```

Data cleaning
```{r}
data <- read.csv("./project_data.csv", header = TRUE)

#removing Booking ID column from our data.
data <- subset(data, select = -c(0,1))

#seeing the length of each room type reserved
table(data$room_type_reserved)
#From the table, we can see that 'room_type1' is largely abundant compared to the other rooms.
#turning this variable into a factor  'room_type1" and 'other
room_type <- data %>%
  group_by(room_type_reserved) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  select(room_type_reserved) %>%
  top_n(-1)
data$room_type_reserved <- ifelse(data$room_type_reserved %in% room_type$room_type_reserved,
                                  data$room_type_reserved,
                                  "other")
unique(data$room_type_reserved)
```
```{r}
unique(data$type_of_meal_plan)
unique(data$market_segment_type)
#I will leave 'type_of_meal_plan' and 'market_segment_type' as is but will still need to convert to a factor
```
```{r}
#Converting dates from "arrival_date" to seasons
data2 <- data
data2 <- data.frame(
  arrival_date = seq(as.Date("2017-01-01"), as.Date("2018-12-31"), by = "days")
)
get_season <- function(month) {
  case_when(
    month %in% 3:5 ~ "Spring",
    month %in% 6:8 ~ "Summer",
    month %in% 9:11 ~ "Fall",
    TRUE ~ "Winter"
  )
}
data2 <- data2 %>%
  mutate(month = month(arrival_date),
         season = get_season(month))
data <- data %>%
  mutate(arrival_date = get_season(month(arrival_date)))

unique(data$arrival_date)
```

Character variables will be changed to factors
```{r}
data$type_of_meal_plan <- factor(data$type_of_meal_plan)
data$room_type_reserved <- factor(data$room_type_reserved)
data$arrival_date <- factor(data$arrival_date)
data$market_segment_type <- factor(data$market_segment_type)
#cancelled is = 1, not_cancelled = 0
data$booking_status <- as.integer(data$booking_status == "canceled")
```

Onehot_encoder
```{r}
training_ind <- createDataPartition(data$booking_status,
                                    p = 0.75,
                                    list = FALSE,
                                    times = 1)
training_set <- data[training_ind,]
test_set <- data[-training_ind,]

onehot_encoder <- dummyVars(~ type_of_meal_plan + room_type_reserved +
                                arrival_date + market_segment_type,
                              training_set[, c("type_of_meal_plan", "room_type_reserved",
                                                 "arrival_date", "market_segment_type")],
                              levelsOnly = TRUE,
                              fullRank = TRUE)

onehot_enc_training <- predict(onehot_encoder,
                                 training_set[, c("type_of_meal_plan", "room_type_reserved",
                                                 "arrival_date", "market_segment_type")])

training_set <- cbind(training_set, onehot_enc_training)

onehot_encoder <- dummyVars(~ type_of_meal_plan + room_type_reserved +
                                arrival_date + market_segment_type,
                              test_set[, c("type_of_meal_plan", "room_type_reserved",
                                                 "arrival_date", "market_segment_type")],
                              levelsOnly = TRUE,
                              fullRank = TRUE)

onehot_enc_test <- predict(onehot_encoder,
                                 test_set[, c("type_of_meal_plan", "room_type_reserved",
                                                 "arrival_date", "market_segment_type")])

test_set <- cbind(test_set, onehot_enc_test)

cols_to_exclude <- c(5,7,9,10,16)

test_set[, -(cols_to_exclude)] <- scale(test_set[, -c(cols_to_exclude)],
                              center = apply(training_set[, -c(cols_to_exclude)], 2, mean),
                              scale = apply(training_set[, -c(cols_to_exclude)], 2, sd))
training_set[, -c(cols_to_exclude)] <- scale(training_set[, -c(cols_to_exclude)])

training_features <- array(data = unlist(training_set[, -c(cols_to_exclude)]),
                                         dim = c(nrow(training_set), 27))
training_labels <- array(data = unlist(training_set[, 16]),
                         dim = c(nrow(training_set)))

test_features <- array(data = unlist(test_set[, -c(cols_to_exclude)]),
                       dim = c(nrow(test_set), 27))
test_labels <- array(data = unlist(test_set[, 16]),
                     dim = c(nrow(test_set)))


```

Loading our environment
```{r}
use_virtualenv("my_tf_workspace")
```

Testing different architectures to see which one will fit the best.

Model_4 (75 units, 50 units, 25 units respectively)
```{r}
model_4 <- keras_model_sequential(list(
  layer_dense(units = 75, activation = "relu"),
  layer_dense(units = 50, activation = "relu"),
  layer_dense(units = 25, activation = "relu"),  
  layer_dense(units = 1, activation = "sigmoid")
))
compile(model_4,
        optimizer = "rmsprop",
        loss = "binary_crossentropy",
        metrics = "accuracy")

early_stopping <- callback_early_stopping(patience = 5)

history_4 <- fit(model_4, training_features, training_labels,
               epochs = 50, batch_size = 512, validation_split = 0.33,
               callbacks = list(early_stopping))
plot(history_4)
```
```{r}
model_5 <- keras_model_sequential(list(
  layer_dense(units = 1000, activation = "relu"),
  layer_dense(units = 500, activation = "relu"),
  layer_dense(units = 250, activation = "relu"),
  layer_dense(units = 100, activation = "relu"),
  layer_dense(units = 1, activation = "sigmoid")
))
compile(model_5,
        optimizer = "rmsprop",
        loss = "binary_crossentropy",
        metrics = "accuracy")

early_stopping <- callback_early_stopping(patience = 5)

history_5 <- fit(model_5, training_features, training_labels,
               epochs = 50, batch_size = 512, validation_split = 0.33,
               callbacks = list(early_stopping))
plot(history_5)
```

```{r}
model_6 <- keras_model_sequential(list(
  layer_dense(units = 1000, activation = "relu"),
  layer_dense(units = 500, activation = "relu"),
  layer_dense(units = 100, activation = "relu"),
  layer_dense(units = 1, activation = "sigmoid")
))
compile(model_6,
        optimizer = "rmsprop",
        loss = "binary_crossentropy",
        metrics = "accuracy")

early_stopping <- callback_early_stopping(patience = 5)

history_6 <- fit(model_6, training_features, training_labels,
               epochs = 50, batch_size = 512, validation_split = 0.33,
               callbacks = list(early_stopping))
plot(history_6)
```
ROC curves for each model

model_4
```{r}
predictions_4 <- predict(model_4, test_features)
test_set_4 <- test_set
test_set_4$p_prob <- predictions_4[,1]
over_threshold <- test_set_4[test_set_4$p_prob >= 0.5,]

roc_data_4 <- data.frame(threshold=seq(1,0,-0.01), fpr=0, tpr=0)
for (i in roc_data_4$threshold) {
  over_threshold <- test_set_4[test_set_4$p_prob >= i, ]
  fpr <- sum(over_threshold$booking_status==0)/sum(test_set_4$booking_status==0)
  roc_data_4[roc_data_4$threshold==i, "fpr"] <- fpr
  
  tpr <- sum(over_threshold$booking_status==1)/sum(test_set_4$booking_status==1)
  roc_data_4[roc_data_4$threshold==i, "tpr"] <- tpr
}

ggplot() +
  geom_line(data = roc_data_4, aes(x = fpr, y = tpr, color = threshold) , linewidth = 2) +
  scale_color_gradientn(colors = rainbow(3)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(data = roc_data_4[seq(1, 101, 10), ], aes(x = fpr, y = tpr)) +
  geom_text(data = roc_data_4[seq(1, 101, 10), ],
            aes(x = fpr, y = tpr, label = threshold, hjust = 1.2 , vjust = -0.2))
```
model_5
```{r}
predictions_5 <- predict(model_5, test_features)
test_set_5 <- test_set
test_set_5$p_prob <- predictions_5[,1]
over_threshold <- test_set_5[test_set_5$p_prob >= 0.5,]

roc_data_5 <- data.frame(threshold=seq(1,0,-0.01), fpr=0, tpr=0)
for (i in roc_data_5$threshold) {
  over_threshold <- test_set_5[test_set_5$p_prob >= i, ]
  fpr <- sum(over_threshold$booking_status==0)/sum(test_set_5$booking_status==0)
  roc_data_5[roc_data_5$threshold==i, "fpr"] <- fpr
  
  tpr <- sum(over_threshold$booking_status==1)/sum(test_set_5$booking_status==1)
  roc_data_5[roc_data_5$threshold==i, "tpr"] <- tpr
}

ggplot() +
  geom_line(data = roc_data_5, aes(x = fpr, y = tpr, color = threshold) , linewidth = 2) +
  scale_color_gradientn(colors = rainbow(3)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(data = roc_data_5[seq(1, 101, 10), ], aes(x = fpr, y = tpr)) +
  geom_text(data = roc_data_5[seq(1, 101, 10), ],
            aes(x = fpr, y = tpr, label = threshold, hjust = 1.2 , vjust = -0.2))
```
model_6
```{r}
predictions_6 <- predict(model_6, test_features)
test_set_6 <- test_set
test_set_6$p_prob <- predictions_6[,1]
over_threshold <- test_set_6[test_set_6$p_prob >= 0.5,]

roc_data_6 <- data.frame(threshold=seq(1,0,-0.01), fpr=0, tpr=0)
for (i in roc_data_6$threshold) {
  over_threshold <- test_set_6[test_set_6$p_prob >= i, ]
  fpr <- sum(over_threshold$booking_status==0)/sum(test_set_6$booking_status==0)
  roc_data_6[roc_data_6$threshold==i, "fpr"] <- fpr
  
  tpr <- sum(over_threshold$booking_status==1)/sum(test_set_6$booking_status==1)
  roc_data_6[roc_data_6$threshold==i, "tpr"] <- tpr
}

ggplot() +
  geom_line(data = roc_data_6, aes(x = fpr, y = tpr, color = threshold) , linewidth = 2) +
  scale_color_gradientn(colors = rainbow(3)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(data = roc_data_6[seq(1, 101, 10), ], aes(x = fpr, y = tpr)) +
  geom_text(data = roc_data_6[seq(1, 101, 10), ],
            aes(x = fpr, y = tpr, label = threshold, hjust = 1.2 , vjust = -0.2))
```
AUC for each model

Model_4 AUC
```{r}
auc_4 <- auc(x = roc_data_4$fpr, y = roc_data_4$tpr, type = "spline")
auc_4
```
Model_5
```{r}
auc_5 <- auc(x = roc_data_5$fpr, y = roc_data_5$tpr, type = "spline")
auc_5
```
Model_6
```{r}
auc_6 <- auc(x = roc_data_6$fpr, y = roc_data_6$tpr, type = "spline")
auc_6
```

Calibration Curves for each model

Model_4
```{r}
calibration_data_4 <- data.frame(bin_midpoint = seq(0.05, 0.95, 0.1),
                                 observed_event_percentage = 0)
for(i in seq(0.05, 0.95, 0.1)) {
  in_interval <- test_set_4[test_set_4$p_prob >= (i-0.05) & test_set_4$p_prob <= (i + 0.05), ]
  oep <- nrow(in_interval[in_interval$booking_status==1, ])/nrow(in_interval)
  calibration_data_4[calibration_data_4$bin_midpoint==i, "observed_event_percentage"] <- oep
}

ggplot(data = calibration_data_4, aes(x = bin_midpoint, y = observed_event_percentage)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(size = 2) +
  geom_text(aes(label = bin_midpoint), hjust = 0.75, vjust = -0.5)
```

model_5
```{r}
calibration_data_5 <- data.frame(bin_midpoint = seq(0.05, 0.95, 0.1),
                                 observed_event_percentage = 0)
for(i in seq(0.05, 0.95, 0.1)) {
  in_interval <- test_set_5[test_set_5$p_prob >= (i-0.05) & test_set_5$p_prob <= (i + 0.05), ]
  oep <- nrow(in_interval[in_interval$booking_status==1, ])/nrow(in_interval)
  calibration_data_5[calibration_data_5$bin_midpoint==i, "observed_event_percentage"] <- oep
}

ggplot(data = calibration_data_5, aes(x = bin_midpoint, y = observed_event_percentage)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(size = 2) +
  geom_text(aes(label = bin_midpoint), hjust = 0.75, vjust = -0.5)
```

model_6
```{r}
calibration_data_6 <- data.frame(bin_midpoint = seq(0.05, 0.95, 0.1),
                                 observed_event_percentage = 0)
for(i in seq(0.05, 0.95, 0.1)) {
  in_interval <- test_set_6[test_set_6$p_prob >= (i-0.05) & test_set_6$p_prob <= (i + 0.05), ]
  oep <- nrow(in_interval[in_interval$booking_status==1, ])/nrow(in_interval)
  calibration_data_6[calibration_data_6$bin_midpoint==i, "observed_event_percentage"] <- oep
}

ggplot(data = calibration_data_6, aes(x = bin_midpoint, y = observed_event_percentage)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point(size = 2) +
  geom_text(aes(label = bin_midpoint), hjust = 0.75, vjust = -0.5)
```